Predictive modeling using genomic annotations

### 2018-09-12 Manuscript/manuscript.Rmd
updates

Due to the large number of genomic features included in the model, different variable selection techniques were evaluated in order to prevent overfitting, improve computational speed, and eliminate uninformative variables. There were three different wrapper methods that were used here, forward, backward, and stepwise selection. A null model consisting of just an intercept term and a full model, including all features, were specified for each selection technique. For all three, a logistic regression model was performed, and the AIC of the fitted model was used to establish which variables to include. That is, only a reduction in the model's AIC contributed to the inclusion of a feature in the model. For each technique, prior to model selection, balanced classes were iteratively created 100 times by random under-sampling from teh majority class. At each iteration the variables selected from each technique were recorded. The final set of predictors selected for each technique were those that appeared over 90% of the time accross all iterations. The union of the three sets of features from each of the techniques was obtained and used as the final feature space of the data (Figure X)

@@

Once the pipeline produced the appropriate dataset, a random forest algorithm was used to assess which genomic features that were most predictive of the development of TAD boundaries. The algorithm was run iteratively over 100 bootstrap samples, each time producing sensitivities, specificities, AUCs, and variable importances. For each iteration, the model was tuned on the best number of predictors to subset as well as the optimal number of trees to use. At the end of the 100 iterations, row wise means were performed to aggregate the model performances. We then compared the random forest model to both the MLR model and the MLR model with LASSO estimated coefficients that was proposed by Mourad et al.

>>>>

Due to the large number of genomic features included in the model, different variable selection techniques were evaluated in order to improve computational speed and eliminate uninformative variables. There were three different wrapper methods that were used here, forward, backward, and stepwise selection. A null model consisting of just an intercept term and a full model, including all features, were specified for each selection technique. For all three, a logistic regression model was performed, and the AIC of the fitted model was used to establish which variables to include. That is, only a reduction in the model's AIC contributed to the inclusion of a feature in the model. For each technique, prior to model selection, balanced classes were iteratively created 100 times by random under-sampling from teh majority class. At each iteration the variables selected from each technique were recorded. The final set of predictors selected for each technique were those that appeared over 90% of the time accross all iterations. The union of the three sets of features from each of the techniques was obtained and used as the final feature space of the data (Figure X)

@@

Once the pipeline produced the appropriate dataset, an elastic was used to assess which genomic features that were most predictive of the development of TAD boundaries. The algorithm was run iteratively over 100 random under-sampled majority classes to balance the data, each time producing performance metrics and variable importances. At the end of the 100 iterations, row-wise means were performed to aggregate the model information. We then compared the elastic-net model performed on the processed data to both a multiple logistic regression model and a LASSO regularized model with and without the use of distance predictors and with or without processed data.

#####################
